% Packaging & Reproducibility (thesis-ready section)
% Include in thesis with: \input{latex/packaging_reproducibility}

\section{Packaging and Reproducibility}
\label{sec:packaging}

All experiments are packaged to enable end-to-end reproducibility of training, evaluation, and plot/table generation. The release contains the full source code, the NPZ datasets used for training/testing, and precomputed artifacts (figures, \LaTeX{} tables, and CSV summaries) for fast inspection.

\subsection{Environment control}
The package includes both \texttt{requirements.txt} (pip) and \texttt{environment.yml} (conda) to recreate the software environment. For maximum reproducibility, users should create the environment first and then execute the provided scripts.

\subsection{Determinism and seeds}
Training scripts accept a \texttt{--seed} argument and support deterministic execution where applicable. This controls Python, NumPy, and PyTorch random generators, enabling stable comparisons across runs.

\subsection{Reproduction modes}
Two reproduction modes are provided:
\begin{itemize}
  \item \textbf{Fast mode} (\texttt{scripts/reproduce\_all\_fast.sh}): regenerates plots and \LaTeX{} tables from the included summaries and cached artifacts.
  \item \textbf{Full mode} (\texttt{scripts/reproduce\_all\_full.sh}): retrains models and reproduces all results from scratch, saving logs, checkpoints, and regenerated figures/tables.
\end{itemize}

\subsection{Outputs}
All generated artifacts are stored under \texttt{outputs/}:
\begin{itemize}
  \item \texttt{outputs/logs/}: training and evaluation logs,
  \item \texttt{outputs/models/}: model checkpoints,
  \item \texttt{outputs/summaries/}: CSV/JSON summaries used to build final tables and plots,
  \item \texttt{outputs/tables/}: \LaTeX{}-ready tables,
  \item \texttt{outputs/figures/}: plots in PDF and PNG formats.
\end{itemize}
